Evaluating Informative Hypotheses With Equality and Inequality Constraint: The Bayes Factor Encompassing Prior Approach

### 2021-11-28 docs/attachment-bayes-factor.tex
update supplemental material

\title{Evaluating Informative Hypotheses With Equality and Inequality Constraint: The Bayes Factor Encompassing Prior Approach}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{Supplemental Material}
\author{\href{https://claudiozandonella.netlify.app/}{Claudio Zandonella Callegher}, Tatiana Marci, Pietro De Carli, and Gianmarco Altoè}
\date{2021-09-21}

@@

This is the Supplemental Material of the article ``\emph{Evaluating Informative Hypotheses With Equality and Inequality Constraint: The Bayes Factor Encompassing Prior Approach}'' {[}TODO: add cit{]}. The aim of the paper is to provide a clear and detailed description of the Bayes Factor with the encompassing prior approach considering an applied example.

@@

When conducting a study, researchers usually have expectations based on hypotheses or theoretical perspectives they want to evaluate. Equality and inequality constraints on the model parameters are used to formalize researchers' expectations or theoretical perspectives into the so-called ``\emph{informative hypotheses}''. However, traditional statistical approaches, such as the Null Hypothesis Significance Testing (NHST) or the model comparison using information criteria (e.g., AIC and BIC), are unsuitable for testing complex informative hypotheses. An alternative approach is to use the Bayes Factor. In particular, the Bayes Factor based on the ``\emph{encompassing prior approach}'' allows researchers to easily evaluate complex informative hypotheses. In this paper we provide a clear and detailed introduction to this approach. We describe all the steps and elements involved in the formalization of informative hypotheses and computation of the Bayes Factor following the encompassing prior approach. Subsequently, we provide an application of this method to a real case scenario comparing the different theoretical perspectives regarding the role of mother attachment and father attachment on children's socio-emotional development.

@@

In the supplemental material we provide the complete statistical analyses of the case study regarding attachment. For more information regarding the Bayes Factor with encompassing prior approach or the formalization of theoretical perspectives regarding the role of mother attachment and father attachment on children socio-emotional development, see the main article {[}TODO: add link article{]}. The supplemental material is structure in three parts:

@@

\hypertarget{exteranlizing-problems}{%
\subsubsection*{2) Exteranlizing Problems}\label{exteranlizing-problems}}
\addcontentsline{toc}{subsubsection}{2) Exteranlizing Problems}

In this section, we present the analyses evaluating the different role of mother attachment and father attachment on children's externalizing problems. We follow three different approaches.

\begin{itemize}
\tightlist
\item
  \textbf{Chapter \ref{model-choice-ext} - Models Family Choice.} We discuss the appropriate models family to to take into account data characteristics.
\item
  \textbf{Chapter \ref{nhst-ext} - NHST.} Analysis are conducted following the traditional Null Hypothesis Significance Testing (NHST).
\item
  \textbf{Chapter \ref{model-comparison-ext} - Model Comparison.} Analysis are conducted according to the Model Comparison approach using the AIC and BIC criteria.
\item
  \textbf{Chapter \ref{BF-ext} - Bayes Factor.} Analysis are conducted following the Bayes Factor with encompassing prior approach.
\item
  \textbf{Chapter \ref{conclusion-ext} - Conclusion.} Results of the three approaches are discussed.
\end{itemize}

\hypertarget{interanlizing-problems}{%
\subsubsection*{3) Interanlizing Problems}\label{interanlizing-problems}}
\addcontentsline{toc}{subsubsection}{3) Interanlizing Problems}

In this section, we present the analyses evaluating the different role of mother attachment and father attachment on children's internalizing problems. We follow the three different approaches as for the externalizing problems analysis. This section is shorter than previous as we focus only on the results without repeating all the descriptions of the methods.

\begin{itemize}
\tightlist
\item
  \textbf{Chapter \ref{model-choice-int} - Models Family Choice.} We discuss the appropriate models family to to take into account data characteristics.
\item
  \textbf{Chapter \ref{nhst-int} - NHST.} Analysis are conducted following the traditional Null Hypothesis Significance Testing (NHST).
\item
  \textbf{Chapter \ref{model-comparison-int} - Model Comparison.} Analysis are conducted according to the Model Comparison approach using the AIC and BIC criteria.
\item
  \textbf{Chapter \ref{BF-int} - Bayes Factor.} Analysis are conducted following the Bayes Factor with encompassing prior approach.
\item
  \textbf{Chapter \ref{conclusion-int} - Conclusion.} Results of the three approaches are discussed.
\end{itemize}

@@

The aim of the study is to compare the different theoretical perspectives regarding the role of mother attachment and father attachment on children's social emotional development. In the literature, four different main theoretical perspectives have been identified \citep{brethertonFathersAttachmentTheory2010}:

\begin{itemize}
\tightlist
\item
  \textbf{Monotropy Theory} - only the mother has an impact on children's development.
\item
  \textbf{Hierarchy Theory} - the mother has a greater impact on children development than the the father.
\item
  \textbf{Independence Theory} - all attachment figures are equally important but they affect the children's development differently.
\item
  \textbf{Integration Theory} - to understand the impact on children's development it is necessary to consider attachment relationships taken together.
\end{itemize}

@@

For more information regarding Attachment theory and the formalization of theoretical perspectives regarding the role of mother attachment and father attachment on children socio-emotional development, see the main article {[}TODO: add link article{]}.

@@

To evaluate the different role of father attachment and mother attachment, we consider 854 Italian children from third to sixth school grade. Only middle-childhood children are included in the analysis (5 participants are excluded as older than 12.3 years and 2 participants are excluded are excluded as the age is missing). The total sample size included in the analysis is 847 (50.65 \% Females). Participants frequencies by grade and gender are reported in Table\textasciitilde\ref{tab:table-gender-grade}.

@@

Participants age mean is 10.28 (SD = 1.34). The participants age summary statistics are reported below and the distribution is presented in Figure\textasciitilde\ref{fig:plot-age}.

@@

Attachment towards the mother and the father was measured separately using the brief Experiences in Close Relationships Scale (ECR; {[}TODO add references{]}) completed by the children. Four main attachment styles have been recognized in the literature according to thee levels of \emph{anxiety} and \emph{avoidance}:

@@

To identify children attachment styles towards the mother and towards the father, we conduct two separate cluster analyses. Clusters are obtained using the function \texttt{hclust()} (with argument \texttt{method="ward.D"}) considering \emph{Euclidean} distances between participants responses to the ECR items.

@@

These results reflect the traditional definition of the four attachment styles. Moreover, attachment styles frequencies are in lines with general results in the literature {[}TODO: add reference values{]}.

@@

To evaluate if 4 cluster is an appropriate choice, we compare the BIC values of different model-based clustering options. To do that we use the function \texttt{mclustBIC()} from the \texttt{mclust} R-package \citep{scruccaMclustClusteringClassification2016}. BIC values of the best three model-based clustering are reported below and overall results are presented in Figure\textasciitilde\ref{fig:plot-mclust-mother}. See \texttt{?mclustBIC()} help page for further information.

@@

These results are no as good as in the mother attachment classification, but they are still reasonable. Moreover, attachment styles frequencies are close to the general results in the literature {[}TODO: add reference values{]}.

@@

To evaluate if 4 cluster is an appropriate choice, we compare the BIC values of different model-based clustering options as before. BIC values of the best three model-based clustering are reported below and overall results are presented in Figure\textasciitilde\ref{fig:plot-mclust-father}.

@@

Children's social-emotional development was measured using the Strength \& Difficulties Questionnaire (SDQ; {[}TODO add references{]}) completed by the teachers. Separate scores for externalizing and internalizing problems were obtained as sum of the questionnaire items.

\hypertarget{externalizing-problems}{%
\subsection{Externalizing Problems}\label{externalizing-problems}}

Participants externalizing problems mean is 3.35 (SD = 3.91). The participants externalizing problems summary statistics are reported below and the distribution is presented in Figure\textasciitilde\ref{fig:plot-externalizing-dist}.

@@

Overall, externalizing problems are low. This is expected as the sample is not a clinical sample. Externalizing problems according to attachment styles are reported in Table\textasciitilde\ref{tab:table-cluster-ext}.

@@

\caption{\label{tab:table-cluster-ext}Externalizing problems according to attachment styles ($n_{subj} = 847$).}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{>{}rcccccccc}
\toprule
\multicolumn{1}{c}{\textbf{ }} & \multicolumn{8}{c}{\textbf{Father Attachment}} \\
\cmidrule(l{3pt}r{3pt}){2-9}
\multicolumn{1}{c}{\textbf{ }} & \multicolumn{2}{c}{\textbf{Secure}} & \multicolumn{2}{c}{\textbf{Anxious}} & \multicolumn{2}{c}{\textbf{Avoidant}} & \multicolumn{2}{c}{\textbf{Fearful}} \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5} \cmidrule(l{3pt}r{3pt}){6-7} \cmidrule(l{3pt}r{3pt}){8-9}
\textbf{Mother Attachemnt} & \textbf{Mean (SD)} & \textbf{Median} & \textbf{Mean (SD)} & \textbf{Median} & \textbf{Mean (SD)} & \textbf{Median} & \textbf{Mean (SD)} & \textbf{Median}\\
\midrule
\textbf{Secure} & 2.63 (3.57) & 1.0 & 3.45 (4.48) & 2.0 & 1.61 (2.13) & 1.0 & 2.88 (3.44) & 2.0\\
\textbf{Anxious} & 3.69 (4.07) & 2.0 & 3.01 (3.61) & 2.0 & 3.32 (4.12) & 2.0 & 4.05 (3.61) & 3.0\\
\textbf{Avoidant} & 2.84 (3.34) & 1.0 & 3.31 (3.65) & 2.0 & 3.71 (4.19) & 2.0 & 3.75 (4.81) & 1.0\\
\textbf{Fearful} & 7.60 (4.04) & 8.0 & 4.64 (3.84) & 4.5 & 4.76 (4.67) & 3.0 & 4.26 (4.07) & 4.0\\
\bottomrule
\end{tabular}}
\end{table}

\hypertarget{internalizing-problems}{%
\subsection{Internalizing Problems}\label{internalizing-problems}}

@@

\begin{verbatim}
##   #
##    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
##   0.000   0.000   2.000   2.955   4.000  18.000
\end{verbatim}

@@

Overall, internalizing problems are low. Again, this is expected as the sample is not a clinical sample. Internalizing problems according to attachment styles are reported in Table\textasciitilde{}\ref(tab:table-cluster-ext).

@@

\caption{\label{tab:table-cluster-int}Internalizing problems according to attachment styles ($n_{subj} = 847$).}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{>{}rcccccccc}
\toprule
\multicolumn{1}{c}{\textbf{ }} & \multicolumn{8}{c}{\textbf{Father Attachment}} \\
\cmidrule(l{3pt}r{3pt}){2-9}
\multicolumn{1}{c}{\textbf{ }} & \multicolumn{2}{c}{\textbf{Secure}} & \multicolumn{2}{c}{\textbf{Anxious}} & \multicolumn{2}{c}{\textbf{Avoidant}} & \multicolumn{2}{c}{\textbf{Fearful}} \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5} \cmidrule(l{3pt}r{3pt}){6-7} \cmidrule(l{3pt}r{3pt}){8-9}
\textbf{Mother Attachemnt} & \textbf{Mean (SD)} & \textbf{Median} & \textbf{Mean (SD)} & \textbf{Median} & \textbf{Mean (SD)} & \textbf{Median} & \textbf{Mean (SD)} & \textbf{Median}\\
\midrule
\textbf{Secure} & 2.29 (2.64) & 1.0 & 2.63 (2.75) & 2.0 & 2.67 (2.75) & 3.0 & 3.62 (3.46) & 3.0\\
\textbf{Anxious} & 3.29 (3.48) & 2.0 & 3.15 (3.09) & 2.0 & 3.28 (3.72) & 2.0 & 3.11 (3.81) & 2.0\\
\textbf{Avoidant} & 1.88 (2.28) & 1.0 & 2.99 (3.22) & 2.0 & 2.65 (2.86) & 2.0 & 4.08 (3.78) & 3.5\\
\textbf{Fearful} & 4.40 (3.65) & 4.0 & 2.07 (1.69) & 2.0 & 3.68 (3.68) & 3.0 & 4.37 (3.12) & 4.0\\
\bottomrule
\end{tabular}}
\end{table}

@@

In this chapter, we discuss the appropriate models family to to take into account data characteristics.

@@

Externalizing problems are computed as the sum of 10 items of the SDQ, obtaining discrete scores that range from 0 to 20. Thus, we should use appropriate discrete distribution such as the \emph{Poisson} distribution or the \emph{Negative Binomial}. In the Poisson distribution mean and variance are defined according to the same parameter \(\lambda\). On the contrary, Negative Binomial has an extra parameter to adjust the variance allowing more flexibility. Considering data distribution (see Figure\textasciitilde\ref{fig:plot-externalizing-dist}), we can observe that data have high dispersion with a long right tail. In this case, Poisson distribution would be a poor choice and we prefer Negative Binomial instead.

Again, considering data distribution (see Figure\textasciitilde\ref{fig:plot-externalizing-dist}), we can observe an high peak of values at zero. Remember that this is not a clinical sample, thus it is expected that children majority has no problems or really few problem. We could question ourselves, however, whether a \emph{Zero-Inflated} model may be appropriate

@@

To evaluate the presence of zero inflation in our data, we compare the number of observed zeros and expected zeros in a Negative Binomial mixed effects model. We consider in the model the role of gender and the interaction between mother attachment and father attachment. Moreover, we consider the children's classroom ID as a random effect to account for teachers' different ability to evaluate children's problems. Using R formula syntax, we have

@@

Results indicate that the model is slightly under-fitting the number of zeros. Now, we can try to fit a \emph{Zero Inflated Negative Binomial} (ZINB) model and compare the performance of the two models. ZINB models are defined as
\[
y_{ij} \sim ZINB(p_{ij}, \mu_{ij}, \phi),
\]
where \(p_{ij}\) is the probability of an observation \(y_{ij}\) being an extra zero (i.e., a zero not coming from the Negative Binomial distribution) and \(1-p_{ij}\) indicates the probability of a given observation \(y_{ij}\) being generated form a Negative Binomial distribution with mean \(\mu_{ij}\) and variance \(\sigma_{ij}^2 = \mu_{ij} + \frac{\mu_{ij}^2}{\phi}\). Moreover, we have that
\[
p_{ij} = \text{logit}^{-1}(X_i^T\beta_p+ Z_j^Tu_p),\\
\mu_{ij} = \text{exp}(X_i^T\beta_{\mu}+ Z_j^Tu_{\mu}).
\]
That is, both \(p\) and \(\mu\) are modelled separately according to (possibly) different variables. In our case, we consider only the role of gender for \(p\) (i.e., the probability of having externalizing problems depends on gender), whereas for \(\mu\) the consider also the interaction between mother attachment and father attachment. In both cases, we consider the children's classroom ID as a random effect (teachers may differ in the ability to detect children's problems and quantify them). Using R formula syntax, we have

@@

The ZINB model is fitted using the function \texttt{zeroinfl()} from the \texttt{pscl} R-package \citep{zeileisRegressionModelsCount2008}. To compare the ZINB model and the Negative Binomial model we conduct an analysis of \emph{Deviance}. Note that, in the case of generalized linear models (GLM), the deviance is the corresponding of the residual variance used in the traditional ANOVA in the case of linear models.

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{emmeans}\SpecialCharTok{::}\FunctionTok{contrast}\NormalTok{(emmeans}\SpecialCharTok{::}\FunctionTok{emmeans}\NormalTok{(fit\_ext\_zinb, }\AttributeTok{specs =} \SpecialCharTok{\textasciitilde{}}\NormalTok{ mother ),}
                  \StringTok{"pairwise"}\NormalTok{, }\AttributeTok{adjust =} \StringTok{"mvt"}\NormalTok{)}
\DocumentationTok{\#\# }\AlertTok{NOTE}\DocumentationTok{: Results may be misleading due to involvement in interactions}
\DocumentationTok{\#\#  contrast           estimate    SE  df t.ratio p.value}
\DocumentationTok{\#\#  Secure {-} Anxious    {-}0.2856 0.140 825  {-}2.034  0.1719}
\DocumentationTok{\#\#  Secure {-} Avoidant   {-}0.3080 0.162 825  {-}1.902  0.2227}
\DocumentationTok{\#\#  Secure {-} Fearful    {-}0.5617 0.178 825  {-}3.149  0.0092}
\DocumentationTok{\#\#  Anxious {-} Avoidant  {-}0.0224 0.124 825  {-}0.181  0.9978}
\DocumentationTok{\#\#  Anxious {-} Fearful   {-}0.2761 0.146 825  {-}1.885  0.2300}
\DocumentationTok{\#\#  Avoidant {-} Fearful  {-}0.2537 0.167 825  {-}1.522  0.4180}
\DocumentationTok{\#\# }
\DocumentationTok{\#\# Results are averaged over the levels of: gender, father }
\DocumentationTok{\#\# Results are given on the log (not the response) scale. }
\DocumentationTok{\#\# P value adjustment: mvt method for 6 tests}
\end{Highlighting}
\end{Shaded}

@@

To interpret these results, note that, AIC tends to select more complex models that can better explain the data, on the contrary, BIC penalizes complex models to a greater extent. As pointed by \citet{kuhaAICBICComparisons2004}, using the two criteria together is always advocated as agreement provides reassurance on the robustness of the results and disagreement still provides useful information for the discussion. We can say that there is evidence in favour of the role of mother attachment but probably this effect is small.

@@

Considering model \texttt{fit\_ext\_mother}, we can run an analysis of deviance to evaluate the significance of the predictors.

@@

\hypertarget{prior-choiche}{%
\subsection{Prior Choiche}\label{prior-choiche}}

>>>>

\title{Evaluating Informative Hypotheses With Equality and Inequality Constraint: The Bayes Factor Encompassing Prior Approach}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{Supplemental Material}
\author{\href{https://claudiozandonella.netlify.app/}{Claudio Zandonella Callegher}, Tatiana Marci, Pietro De Carli, and Gianmarco Altoè}
\date{2021-11-28}

@@

This is the Supplemental Material of the article ``\emph{Evaluating informative hypotheses with equality and inequality constraints: A tutorial using the Bayes factor via the encompassing prior approach}'' {[}TODO: add cit{]}. The paper aims to provide a clear and detailed description of the Bayes Factor with the encompassing prior approach considering an applied example.

@@

When conducting a study, researchers usually have expectations based on hypotheses or theoretical perspectives they want to evaluate. Equality and inequality constraints on the model parameters are used to formalize researchers' expectations or theoretical perspectives into the so-called informative hypotheses. However, traditional statistical approaches, such as the Null Hypothesis Significance Testing (NHST) or the model comparison using information criteria (e.g., AIC and BIC), are unsuitable for testing complex informative hypotheses. An alternative approach is to use the Bayes factor. In particular, the Bayes factor based on the encompassing prior approach allows researchers to easily evaluate complex informative hypotheses in a wide range of statistical models (e.g., generalized linear). This paper provides a detailed introduction to the Bayes factor with encompassing prior. First, all steps and elements involved in the formalization of informative hypotheses and the computation of the Bayes factor with encompassing prior are described. Next, we apply this method to a real case scenario, considering the attachment theory. Specifically, we analyzed the relative influence of maternal and paternal attachment on children's social-emotional development by comparing the various theoretical perspectives debated in the literature.

@@

In the supplemental material, we provide the complete statistical analyses of the case study regarding attachment. For more information regarding the Bayes factor with encompassing prior approach or the formalization of theoretical perspectives regarding the role of mother attachment and father attachment on children socio-emotional development, see the main article {[}TODO: add link article{]}. The supplemental material is structured into three parts:

@@

\hypertarget{externalizing-problems}{%
\subsubsection*{2) Externalizing Problems}\label{externalizing-problems}}
\addcontentsline{toc}{subsubsection}{2) Externalizing Problems}

In this section, we present the analyses evaluating the different roles of mother attachment and father attachment on children's externalizing problems. We follow three different approaches.

\begin{itemize}
\tightlist
\item
  \textbf{Chapter \ref{model-choice-ext} - Models Family Choice.} We discuss the appropriate models' family to take into account data characteristics.
\item
  \textbf{Chapter \ref{nhst-ext} - NHST.} Analysis is conducted following the traditional Null Hypothesis Significance Testing (NHST).
\item
  \textbf{Chapter \ref{model-comparison-ext} - Model Comparison.} Analysis is conducted according to the Model Comparison approach using the AIC and BIC criteria.
\item
  \textbf{Chapter \ref{BF-ext} - Bayes Factor.} Analysis is conducted following the Bayes Factor with the encompassing prior approach.
\item
  \textbf{Chapter \ref{conclusion-ext} - Conclusions.} Results of the three approaches are discussed.
\end{itemize}

\hypertarget{internalizing-problems}{%
\subsubsection*{3) Internalizing Problems}\label{internalizing-problems}}
\addcontentsline{toc}{subsubsection}{3) Internalizing Problems}

In this section, we present the analyses evaluating the different roles of mother attachment and father attachment on children's internalizing problems. We follow the three different approaches like in the analysis of the externalizing problems. This section is shorter than the previous one as we focus only on the results without repeating all the descriptions of the methods.

\begin{itemize}
\tightlist
\item
  \textbf{Chapter \ref{model-choice-int} - Models Family Choice.} We discuss the appropriate models' family to take into account data characteristics.
\item
  \textbf{Chapter \ref{nhst-int} - NHST.} Analysis is conducted following the traditional Null Hypothesis Significance Testing (NHST).
\item
  \textbf{Chapter \ref{model-comparison-int} - Model Comparison.} Analysis is conducted according to the Model Comparison approach using the AIC and BIC criteria.
\item
  \textbf{Chapter \ref{BF-int} - Bayes Factor.} Analysis is conducted following the Bayes Factor with the encompassing prior approach.
\item
  \textbf{Chapter \ref{conclusion-int} - Conclusions.} Results of the three approaches are discussed.
\end{itemize}

@@

The study aims to compare the different theoretical perspectives regarding the role of mother attachment and father attachment on children's social-emotional development. In the literature, four different main theoretical perspectives have been identified \citep{brethertonFathersAttachmentTheory2010}:

\begin{itemize}
\tightlist
\item
  \textbf{Monotropy Theory} - only the mother has an impact on children's development.
\item
  \textbf{Hierarchy Theory} - the mother has a greater impact on children's development than the father.
\item
  \textbf{Independence Theory} - all attachment figures are equally important but they affect the children's development differently.
\item
  \textbf{Integration Theory} - to understand the impact on children's development it is necessary to consider attachment relationships taken together.
\end{itemize}

@@

For more information regarding Attachment theory and the formalization of theoretical perspectives regarding the role of mother attachment and father attachment on children's socio-emotional development, see the main article {[}TODO: add link article{]}.

@@

To evaluate the different roles of father attachment and mother attachment, we consider 854 Italian children from third to sixth school grade. Only middle-childhood children are included in the analysis (5 participants are excluded as older than 12.3 years and 2 participants are excluded as the age is missing). The total sample size included in the analysis is 847 (50.65 \% Females). Participants' frequencies by grade and gender are reported in Table\textasciitilde\ref{tab:table-gender-grade}.

@@

Participants' age mean is 10.28 (SD = 1.34). The participants' age summary statistics are reported below and the distribution is presented in Figure\textasciitilde\ref{fig:plot-age}.

@@

Attachment towards the mother and the father was measured separately using the brief Experiences in Close Relationships Scale - Revised Child Version {[}ECR-RC, \citet{Brenning2014ThePQ}; marciBriefExperiencesClose2019{]} completed by the children. Four main attachment styles have been recognized in the literature according to thee levels of \emph{anxiety} and \emph{avoidance}:

@@

To identify children's attachment styles towards the mother and towards the father, we conduct two separate cluster analyses. Clusters are obtained using the function \texttt{hclust()} (with argument \texttt{method="ward.D"}) considering \emph{Euclidean} distances between participants responses to the ECR items.

@@

These results reflect the traditional definition of the four attachment styles.

@@

To evaluate if 4 clusters is an appropriate choice, we compare the BIC values of different model-based clustering options. To do that we use the function \texttt{mclustBIC()} from the \texttt{mclust} R-package \citep{scruccaMclustClusteringClassification2016}. BIC values of the best three model-based clusterings are reported below and overall results are presented in Figure\textasciitilde\ref{fig:plot-mclust-mother}. See \texttt{?mclustBIC()} help page for further information.

@@

These results are not as good as in the mother attachment classification, but they are still reasonable.

@@

To evaluate if 4 clusters is an appropriate choice, we compare the BIC values of different model-based clustering options as before. BIC values of the best three model-based clusterings are reported below and overall results are presented in Figure\textasciitilde\ref{fig:plot-mclust-father}.

@@

Children's social-emotional development was measured using the Strength and Difficulties Questionnaire \citep[SDQ,][]{goodmanWhenUseBroader2010} completed by the teachers. Separate scores for externalizing and internalizing problems were obtained as the sum of the questionnaire items.

\hypertarget{externalizing-problems-1}{%
\subsection{Externalizing Problems}\label{externalizing-problems-1}}

Participants externalizing problems mean is 3.35 (SD = 3.91). The participants' externalizing problems summary statistics are reported below and the distribution is presented in Figure\textasciitilde\ref{fig:plot-externalizing-dist}.

@@

Overall, externalizing problems are low. This is expected as the sample is not clinical. Externalizing problems according to attachment styles are reported in Table\textasciitilde\ref{tab:table-cluster-ext}.

@@

\caption{\label{tab:table-cluster-ext}Externalizing problems according to attachment styles ($n_{subj} = 847$).}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{>{}rcccccccc}
\toprule
\multicolumn{1}{c}{\textbf{ }} & \multicolumn{8}{c}{\textbf{Father Attachment}} \\
\cmidrule(l{3pt}r{3pt}){2-9}
\multicolumn{1}{c}{\textbf{ }} & \multicolumn{2}{c}{\textbf{Secure}} & \multicolumn{2}{c}{\textbf{Anxious}} & \multicolumn{2}{c}{\textbf{Avoidant}} & \multicolumn{2}{c}{\textbf{Fearful}} \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5} \cmidrule(l{3pt}r{3pt}){6-7} \cmidrule(l{3pt}r{3pt}){8-9}
\textbf{Mother Attachment} & \textbf{Mean (SD)} & \textbf{Median} & \textbf{Mean (SD)} & \textbf{Median} & \textbf{Mean (SD)} & \textbf{Median} & \textbf{Mean (SD)} & \textbf{Median}\\
\midrule
\textbf{Secure} & 2.63 (3.57) & 1.0 & 3.45 (4.48) & 2.0 & 1.61 (2.13) & 1.0 & 2.88 (3.44) & 2.0\\
\textbf{Anxious} & 3.69 (4.07) & 2.0 & 3.01 (3.61) & 2.0 & 3.32 (4.12) & 2.0 & 4.05 (3.61) & 3.0\\
\textbf{Avoidant} & 2.84 (3.34) & 1.0 & 3.31 (3.65) & 2.0 & 3.71 (4.19) & 2.0 & 3.75 (4.81) & 1.0\\
\textbf{Fearful} & 7.60 (4.04) & 8.0 & 4.64 (3.84) & 4.5 & 4.76 (4.67) & 3.0 & 4.26 (4.07) & 4.0\\
\bottomrule
\end{tabular}}
\end{table}

\hypertarget{internalizing-problems-1}{%
\subsection{Internalizing Problems}\label{internalizing-problems-1}}

@@

\begin{verbatim}
##   # Internalizing problems summary statistics
##    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
##   0.000   0.000   2.000   2.955   4.000  18.000
\end{verbatim}

@@

Overall, internalizing problems are low. Again, this is expected as the sample is not clinical. Internalizing problems according to attachment styles are reported in Table\textasciitilde{}\ref(tab:table-cluster-ext).

@@

\caption{\label{tab:table-cluster-int}Internalizing problems according to attachment styles ($n_{subj} = 847$).}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{>{}rcccccccc}
\toprule
\multicolumn{1}{c}{\textbf{ }} & \multicolumn{8}{c}{\textbf{Father Attachment}} \\
\cmidrule(l{3pt}r{3pt}){2-9}
\multicolumn{1}{c}{\textbf{ }} & \multicolumn{2}{c}{\textbf{Secure}} & \multicolumn{2}{c}{\textbf{Anxious}} & \multicolumn{2}{c}{\textbf{Avoidant}} & \multicolumn{2}{c}{\textbf{Fearful}} \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5} \cmidrule(l{3pt}r{3pt}){6-7} \cmidrule(l{3pt}r{3pt}){8-9}
\textbf{Mother Attachment} & \textbf{Mean (SD)} & \textbf{Median} & \textbf{Mean (SD)} & \textbf{Median} & \textbf{Mean (SD)} & \textbf{Median} & \textbf{Mean (SD)} & \textbf{Median}\\
\midrule
\textbf{Secure} & 2.29 (2.64) & 1.0 & 2.63 (2.75) & 2.0 & 2.67 (2.75) & 3.0 & 3.62 (3.46) & 3.0\\
\textbf{Anxious} & 3.29 (3.48) & 2.0 & 3.15 (3.09) & 2.0 & 3.28 (3.72) & 2.0 & 3.11 (3.81) & 2.0\\
\textbf{Avoidant} & 1.88 (2.28) & 1.0 & 2.99 (3.22) & 2.0 & 2.65 (2.86) & 2.0 & 4.08 (3.78) & 3.5\\
\textbf{Fearful} & 4.40 (3.65) & 4.0 & 2.07 (1.69) & 2.0 & 3.68 (3.68) & 3.0 & 4.37 (3.12) & 4.0\\
\bottomrule
\end{tabular}}
\end{table}

@@

In this chapter, we discuss the appropriate models' family to take into account data characteristics.

@@

Externalizing problems are computed as the sum of 10 items of the SDQ, obtaining discrete scores that range from 0 to 20. Thus, we should use appropriate discrete distribution such as the \emph{Poisson} distribution or the \emph{Negative Binomial}. In the Poisson distribution mean and variance are defined according to the same parameter \(\lambda\). On the contrary, Negative Binomial has an extra parameter to adjust the variance allowing more flexibility. Considering data distribution (see Figure\textasciitilde\ref{fig:plot-externalizing-dist}), we can observe that data have high dispersion with a long right tail. In this case, the Poisson distribution would be a poor choice and we prefer Negative Binomial instead.

Again, considering data distribution (see Figure\textasciitilde\ref{fig:plot-externalizing-dist}), we can observe a high peak of values at zero. Remember that this is not a clinical sample, thus it is expected that the majority of children have no problems or really few problems. We could question ourselves, however, whether a \emph{Zero-Inflated} model may be appropriate

@@

To evaluate the presence of zero inflation in our data, we compare the number of observed zeros and expected zeros in a Negative Binomial mixed-effects model. We consider in the model the role of gender and the interaction between mother attachment and father attachment. Moreover, we consider the children's classroom ID as a random effect to account for teachers' different ability to evaluate children's problems. Using R formula syntax, we have

@@

Results indicate that the model is slightly under-fitting the number of zeros. Now, we can try to fit a \emph{Zero Inflated Negative Binomial} (ZINB) model and compare the performance of the two models. ZINB models are defined as
\[
y_{ij} \sim ZINB(p_{ij}, \mu_{ij}, \phi),
\]
where \(p_{ij}\) is the probability of an observation \(y_{ij}\) being an extra zero (i.e., a zero not coming from the Negative Binomial distribution) and \(1-p_{ij}\) indicates the probability of a given observation \(y_{ij}\) being generated form a Negative Binomial distribution with mean \(\mu_{ij}\) and variance \(\sigma_{ij}^2 = \mu_{ij} + \frac{\mu_{ij}^2}{\phi}\). Moreover, we have that
\[
p_{ij} = \text{logit}^{-1}(X_i^T\beta_p+ Z_j^Tu_p),\\
\mu_{ij} = \text{exp}(X_i^T\beta_{\mu}+ Z_j^Tu_{\mu}).
\]
That is, both \(p\) and \(\mu\) are modelled separately according to (possibly) different variables. In our case, we consider only the role of gender for \(p\) (i.e., the probability of having externalizing problems depends on gender), whereas for \(\mu\) we also consider the interaction between mother attachment and father attachment. In both cases, we consider the children's classroom ID as a random effect (teachers may differ in the ability to detect children's problems and quantify them). Using R formula syntax, we have

@@

The ZINB model is fitted using the function \texttt{glmmTMB()}. To compare the ZINB model and the Negative Binomial model we conduct an analysis of \emph{Deviance}. Note that, in the case of generalized linear models (GLM), the deviance is the corresponding of the residual variance used in the traditional ANOVA in the case of linear models.

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{emmeans}\SpecialCharTok{::}\FunctionTok{contrast}\NormalTok{(emmeans}\SpecialCharTok{::}\FunctionTok{emmeans}\NormalTok{(fit\_ext\_zinb, }\AttributeTok{specs =} \SpecialCharTok{\textasciitilde{}}\NormalTok{ mother ),}
                  \StringTok{"pairwise"}\NormalTok{, }\AttributeTok{adjust =} \StringTok{"mvt"}\NormalTok{)}
\DocumentationTok{\#\# }\AlertTok{NOTE}\DocumentationTok{: Results may be misleading due to involvement in interactions}
\DocumentationTok{\#\#  contrast           estimate    SE  df t.ratio p.value}
\DocumentationTok{\#\#  Secure {-} Anxious    {-}0.2856 0.140 825  {-}2.034  0.1720}
\DocumentationTok{\#\#  Secure {-} Avoidant   {-}0.3080 0.162 825  {-}1.902  0.2228}
\DocumentationTok{\#\#  Secure {-} Fearful    {-}0.5617 0.178 825  {-}3.149  0.0089}
\DocumentationTok{\#\#  Anxious {-} Avoidant  {-}0.0224 0.124 825  {-}0.181  0.9978}
\DocumentationTok{\#\#  Anxious {-} Fearful   {-}0.2761 0.146 825  {-}1.885  0.2300}
\DocumentationTok{\#\#  Avoidant {-} Fearful  {-}0.2537 0.167 825  {-}1.522  0.4180}
\DocumentationTok{\#\# }
\DocumentationTok{\#\# Results are averaged over the levels of: gender, father }
\DocumentationTok{\#\# Results are given on the log (not the response) scale. }
\DocumentationTok{\#\# P value adjustment: mvt method for 6 tests}
\end{Highlighting}
\end{Shaded}

@@

To interpret these results, note that, AIC tends to select more complex models that can better explain the data, on the contrary, BIC penalizes complex models to a greater extent. As pointed out by \citet{kuhaAICBICComparisons2004}, using the two criteria together is always advocated as agreement provides reassurance on the robustness of the results and disagreement still provides useful information for the discussion. We can say that there is evidence in favour of the role of mother attachment but probably this effect is small.

@@

Considering the model \texttt{fit\_ext\_mother}, we can run an analysis of deviance to evaluate the significance of the predictors.

@@

\hypertarget{prior-choice}{%
\subsection{Prior Choice}\label{prior-choice}}
